%\TUD@setbigtitlestyle{\section}{\sectfont\chapterlinesize}{\thechapter}{\TUD@chapind}{-1}
\chapter*{}

\textbf{Er\-kl\char228\-rung zur Abschlussarbeit gem\char228\ss{} \S\,23 Abs. 7 APB der TU Darmstadt}

%\parindent0em
\vskip4ex
Hiermit versichere ich, Sebastian Rinder, die vorliegende Bachelor-Thesis ohne Hilfe Dritter und nur mit den an\-ge\-ge\-benen Quellen und
Hilfsmitteln angefertigt zu haben. Alle Stellen, die Quellen entnommen wurden, sind als solche kenntlich gemacht worden. Diese Arbeit hat in gleicher oder \char228 hn\-lich\-er Form noch keiner Pr\char252\-fungs\-be\-h\char246 r\-de vorgelegen.

\vskip2ex Mir ist bekannt, dass im Fall eines Plagiats (\S38 Abs. 2 APB) ein T\char228uschungsversuch vorliegt,
der dazu f\char252hrt, dass die Arbeit mit 5,0 bewertet und damit ein Pr\char252fungsversuch verbraucht wird.
Abschlussarbeiten d\char252rfen nur einmal wiederholt werden.

\vskip2ex Bei der abgegebenen Thesis stimmen die schriftliche und die zur Archivierung eingereichte elektronische Fassung
\char252berein.
\vskip4ex
\rule{\textwidth}{0.5pt}
\vskip4ex
\textbf{English translation for information purposes only:}
\vskip2ex
\textbf{Thesis Statement pursuant to ยง 23 paragraph 7 of APB TU Darmstadt}
%\parindent0em
\vskip4ex

I herewith formally declare that I, Sebastian Rinder, have written the submitted thesis
independently. I did not use any outside support except for the quoted literature and other
sources mentioned in the paper. I clearly marked and separately listed all of the literature and
all of the other sources which I employed when producing this academic work, either literally
or in content. This thesis has not been handed in or published before in the same or similar
form.

\vskip2ex I am aware, that in case of an attempt at deception based on plagiarism (ยง38 Abs. 2 APB), the
thesis would be graded with 5,0 and counted as one failed examination attempt. The thesis
may only be repeated once.

\vskip2ex In the submitted thesis the written copies and the electronic version for archiving are identical
in content.

\vskip6ex
\begin{tabular}{lp{1cm}l}
Datum / Date: & & Unterschrift / Signature:\\[4ex]
\tudrule[0.3\textwidth] & & \tudrule[0.5\textwidth] \end{tabular}

\if@TUD@resttwocol\twocolumn\fi%
\clearpage
